---
title: "TaxNorm Introduction"
output: 
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{TaxNorm Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

This document introduces the `TaxNorm` R package, a package for normalizing microbiome taxa data. Here, we will go through how to install, analyze and visualize microbiome data using this package. `TaxNorm` implements the Zero Inflated Negative Binomial (ZINB) method to normalize microbiome data. 

## What is the ZINB method?


## Outline

There are three main steps in using this package: 

- **Load and QC Input Data**: In the package we have an example data set from the phyloeq package that shows shows the format needed for analysis. These data can be generated using methods blah blah blah. 

- **Running ZINB Normalization Function**: The `TaxNorm_Normalization` function is runn using the above data on the input. This function implements the ZINB method for normalization.

- **Visualizing and Quality Control**: Last, visualization and quality control measures are built into the package for use. 


## Installation

### Required Packages 

`TaxNorm` requires the packages `phyloeq` and `microbiome` which can be found on bioconductor. 

### Installation from Bioconductor

### Installation from Github

For the newest, but potentially unstable, version of the package, direct github installation is also supported.

```{r, eval = FALSE}
remotes::install_github("wangziyue57/TaxNorm")
```

### Loading Package into R Environment 

```{r, eval = FALSE}
library(TaxNorm)
# library(phyloseq)
# library(microbiome)
# library(ggplot2)
# library(vegan)
# library(MASS)
```



## Example Usage

Basic Useage
```{r, eval = FALSE}
data("TaxNorm_Example_Input", package = "TaxNorm")

# run normalization
TaxNorm_Example_Output <- TaxNorm_Normalization(data= TaxNorm_Example_Input, 
                                         depth = NULL, 
                                         group = sample_data(TaxNorm_Example_Input)$body_site, 
                                         meta.data = NULL,
                                  filter.cell.num = 10,
                                  filter.taxa.count = 0,
                                  random = FALSE,
                                  ncores = NULL)

#save(TaxNorm_Example_Output, file = "../data/TaxNorm_Example_Output.rda")
# run diagnosis test
Diagnose_Data <- TaxNorm_Run_Diagnose(Normalized_Results = TaxNorm_Example_Output, prev = TRUE, equiv = TRUE, group = sample_data(TaxNorm_Example_Input)$body_site)

```


### Load Input Data

Built in example data as a phyloseq object can be loaded with the command below. 

```{r, eval = FALSE}
data("TaxNorm_Example_Input", package = "TaxNorm")

```


### Pre-process Input Data

We have prepared several QC figures for the input data characters, which give a preliminary criteria on pre-filtering rare taxa with low information before any analysis. This will improve the power and computational efficiency for the algorithm. If the user already has the cleaned data or pre-processed the data by themselves before, they can ignore and skip this step.

```{r, eval = FALSE}
qc_data <- TaxNorm_QC_Input(TaxNorm_Example_Input)
```

Here we provide a popular option to ensure at least `filter.sample.num` samples with a count of `filter.taxa.count` or more, where `filter.sample.num` can be chosen as any arbitrary value or the sample size of the smallest group of samples. By default, we used `filter.taxa.count=1` and `filter.sample.num=10`. This criteria is incorporated in the following main function `TaxNorm_Normalization()` as well. 

```{r, eval = FALSE}
filter.sample.num =1
filter.taxa.count = 10
taxaIn <- rowSums(abundances(TaxNorm_Example_Input) > filter.taxa.count) > filter.sample.num
TaxNorm_Example_Input <- prune_taxa(taxaIn, TaxNorm_Example_Input) 
```

Users can apply any of their customized filtering criteria as well. Alternatively, a basic pre-filtering is to keep only rows that have at least 10 reads total:

```{r, eval = FALSE}
taxaIn <- rowSums(abundances(TaxNorm_Example_Input)) > 10
TaxNorm_Example_Input <- prune_taxa(taxaIn, TaxNorm_Example_Input) 
```

### QC Input Data

```{r, eval = FALSE}
qc_data <- TaxNorm_QC_Input(TaxNorm_Example_Input)
```

### Run Normalization

The normalization is run and returns a `TaxNorm_Results` object. This object contains the input data, raw data, normdata, ecdf, model parameters, and convergence. 

```{r, eval = FALSE}
#Pick group from phyloseq object
group <- sample_data(TaxNorm_Example_Input)$body_site
#Run Normalization function
Normalized_Data <- TaxNorm_Normalization(data = TaxNorm_Example_Input, 
                     depth = NULL,
                     group = group, 
                     filter.taxa.count = 0,
                     random = TRUE,
                     ncores = 1)

```



### QC TaxNorm Model

```{r, eval = FALSE}

data("TaxNorm_Example_Output", package = "TaxNorm")


TaxNorm_Model_QC(TaxNormResults = TaxNorm_Example_Output)
```


### TaxNorm NMDS
# Not running currently 

```{r, eval = FALSE}

TaxNorm_NMDS(TaxNormResults = TaxNorm_Example_Output, group_column = "body_site")

```



