---
title: "TaxNorm Introduction"
output: 
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{TaxNorm Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

This document introduces the `TaxNorm` R package, a package for normalizing microbiome taxa data. Here, we will go through how to install, analyze and visualize microbiome data using this package. `TaxNorm` implements the Zero Inflated Negative Binomial (ZINB) method to normalize microbiome data. 

## What is the ZINB method?


## Outline

There are three main steps in using this package: 

- **Load and QC Input Data**: In the package we have an example data set from the phyloeq package that shows shows the format needed for analysis. These data can be generated using methods blah blah blah. 

- **Running ZINB Normalization Function**: The `TaxNorm_Normalization` function is runn using the above data on the input. This function implements the ZINB method for normalization.

- **Visualizing and Quality Control**: Last, visualization and quality control measures are built into the package for use. 


## Installation

### Required Packages 

`TaxNorm` requires the packages `phyloeq` and `microbiome` which can be found on bioconductor. 

### Installation from Bioconductor

### Installation from Github

For the newest, but potentially unstable, version of the package, direct github installation is also supported.

```{r, eval = FALSE}
remotes::install_github("wangziyue57/TaxNorm")
```

### Loading Package into R Environment 

```{r, eval = FALSE}
library(TaxNorm)
library(phyloseq)
library(ggplot2)
library(vegan)
```



## Example Usage


### Load Input Data

Built in example data can be loaded with the command below. Data must be a phyloseq object or a count matrix. 

```{r, eval = FALSE}

data("TaxNorm_Example_Input", package = "TaxNorm")


```

### Convert Count Matrix to Phyloseq

```{r, eval = FALSE}

```


### QC Input Data

```{r, eval = FALSE}
qc_data <- TaxNorm_QC_Input(TaxNorm_Example_Input)
```

### Run Normalization

The normalization is run and returns a `TaxNorm_Results` object. This object contains the input data, raw data, normdata, ecdf, model parameters, and convergence. 

```{r, eval = FALSE}
#Pick group from phyloseq object
group <- sample_data(TaxNorm_Example_Input)$body_site
#Run Normalization function
Normalized_Data <- TaxNorm_Normalization(data = TaxNorm_Example_Input, 
                     depth = NULL,
                     group = group, 
                     filter.sample.num = 1,
                     filter.taxa.count = 0,
                     random = TRUE,
                     ncores = 1)

```



### QC TaxNorm Model

```{r, eval = FALSE}

data("TaxNorm_Example_Output", package = "TaxNorm")


TaxNorm_Model_QC(TaxNormResults = TaxNorm_Example_Output)
```


### TaxNorm NMDS

```{r, eval = FALSE}

TaxNorm_NMDS(TaxNorm_Example_Output, group = "body_site")

```



