---
title: "TaxNorm Introduction"
output: 
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{toxpiR Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<style type="text/css">

h1.title {
  font-size: 28px;
}
h1 { /* Header 1 */
  font-size: 20px;
}
h2 { /* Header 2 */
    font-size: 20px;
}
h3 { /* Header 3 */
  font-size: 18px;
}
h4 { /* Header 4 */
  font-size: 16px;
   color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
</style>


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

This document introduces the `TaxNorm` R package, a package for normalizing microbiome taxa data. Here, we will go through how to install, analyze and visualize microbiome data using this package. `TaxNorm` implements the Zero Inflated Negative Binomial (ZINB) method to normalize microbiome data. 

## What is the ZINB method?


## Outline

There are three main steps in using this package: 

- **Loading Analysis Ready Data**: In the package we have an example data set from the phyloeq package that shows shows the format needed for analysis. These data can be generated using methods blah blah blah. 

- **Running ZINB Normalization Function**: The `TaxNorm_Normalization` function is runn using the above data on the input. This function impliments the ZINB method for normalization.

- **Visualizing and Quality Control**: Last, visualization and quality control measures are built into the package for use. 


## Installation

### Required Packages 

`TaxNorm` requires the packages `phyloeq` and `microbiome` which can be found on bioconductor. 

### Installation from Bioconductor

### Installation from Github

For the newest, but potentially unstable, version of the package, direct github installation is also supported.

```{r, eval = FALSE}
remotes::install_github("wangziyue57/TaxNorm")
```

### Loading Package into R Environment 

```{r, eval = FALSE}
library(TaxNorm)
```



## Example Usage










